---
title: "rpapoda Phase 1: Setting the Stage"
output: html_notebook
bibliography: bib.bib
---



### For each variable:  
* Units  
* Levels of factor variables
* Information about any missing data

```{r variable names}
names(ks)
```

```{r}
head(ks)
```


### Variables
<!-- I think its best to wait until the end to do this part. Too much will change between now and then -->

`r colnames(ks[1])`: A unique number assigned by Kickstarter to identify each project  
`r colnames(ks[2])`: The name of each project  
`r colnames(ks[3])`: A short description of each project  
`r colnames(ks[4])`: The target goal to be raised for each project. Units:  
`r colnames(ks[5])`: The amount pledged by backers to the project. Units: 
`r colnames(ks[6])`: The current (at the time of download) state the project is in. Levels: `r levels(ks$state)`  
`r colnames(ks[7])`: 
`r colnames(ks[8])`: 
`r colnames(ks[9])`: 
`r colnames(ks[10])`: 
`r colnames(ks[11])`: 
`r colnames(ks[12])`: 
`r colnames(ks[13])`: 
`r colnames(ks[14])`: 
`r colnames(ks[15])`: 
`r colnames(ks[16])`: 
`r colnames(ks[17])`: 
`r colnames(ks[18])`: 
`r colnames(ks[19])`: 
`r colnames(ks[20])`: 
`r colnames(ks[21])`: 
`r colnames(ks[22])`: 
`r colnames(ks[23])`: 
`r colnames(ks[24])`: 
`r colnames(ks[25])`: 
`r colnames(ks[26])`: 

