---
title: "Quick Reference"
output: html_notebook
---
## Iteration Cycle
My goal based iteration cycle is currently as follows:
1. Set a goal for the task ahead
* Define expectations or criteria for success
2. Attempt to complete the goal or perform the task
3. Compare your results to your expectations or your criteria for success
4. (Optional) Use judgment and experience to decide whether you think the expectations should be altered, or the task should be repeated under different parameters
5. Make adjustments, and repeat as necessary

**My Process of EDA**
<!-- Should be customized a bit -->
From Signal and NUCI
1. Variation within categories: Bar charts, dot plots (Ranking and part-to-whole relationships)
2. Variation within measures: Histograms, frequency polygons, boxplots (single and multiple distributions, aka distributions by category) # Distributions
3. Variation through time: Line graphs*
4. Relationships among measures: Scatterplots, scatterplot matrices # Correlations
5. Relationships among categories: Visual crosstabs/small multiples with horizontal bar charts or scatterplots 
<!-- Could this replace the goals above as well? -->

Variation within categories:
The Categories:
state
country
category
staff_pick
spotlight
creator

The Variation:
state
mean/median pledged_to_goal
usd_pledged
size of the goal 
backer_count
average pledged

## Generic Process for Each Graph 
<!-- Works well for Variation within Categories -->
Before the graph
Goal, question to be answered, or insight sought: 
Type of graph: 
Variables to be used: 
Geoms or layers to be used: 
```{r example of graphing process}

```

After the graph
Insights gained: 
Additional questions raised: 
Adjustments to be made (if any): 

## Variation Within Measures
Distributions to Explore:
usd_pledged
pledged_to_goal
usd_goal
backers_count

Categories to group by:
state
country
staff_pick
category

Vocab for assessing the shape of a distribution:
curved
flat
upward
downward
single peaked
bi-modal
symmetrical
skewed
concentrations
gaps

## SOCS
shape
outliers
center
spread


## Process for Each Distribution Analysis
Before the graph
Distribution(s) to be analysed/compared: 
Goal, question to be answered, or insight sought: 
What I know about or expect from the distribution(s), including shape, outliers, center, spread: 
Additional variables to be used: 
Geom(s) to be used: 
```{r distribution analysis}



```
After the graph
How does the visualization compare to my expectations, including shape, outliers, center, spread: 
Insights gained: 
Additional questions raised: 
Adjustments to be made (if any): 

## Relationships Among Measures
usd_pledged (likely dependent/response)
pledged_to_goal (likely dependent/response)
usd_goal
backers_count
mean_pledge

Categories:
state (essentially a response)
country 
staff_pick
category

So we need 2 continuous variables, which we can also separate into categories, and map another continuous variable by size.
Pairs
usd_pledged ~ usd_goal
usd_pledged ~ usd_goal, by staff_pick
usd_pledged ~ usd_goal, size = mean_pledge
usd_pledged ~ backers_count
usd_pledged ~ backers_count, by staff_pick
pledged_to_goal ~ usd_goal
pledged_to_goal ~ usd_goal, by category
pledged_to_goal ~ usd_goal, by staff_pick
pledged_to_goal ~ backers_count
pledged_to_goal ~ backers_count, by staff_pick

Facetting:
facet_wrap:
by state
by staff_pick
category?

facet_grid:
state ~ staff_pick


## Process for Each Correlation Analysis
Before the graph
Relationship(s) to be analysed/compared: 
Goal, question to be answered, or insight sought: 
What I know about or expect from the relationship(s), including direction, strength, and shape: 
Do I expect concentrations, gaps, or outliers: 
Additional variables to be used: 
Geom(s) to be used: 
```{r correlation analysis}


```
After the graph
How does the visualization compare to my expectations, including direction, strength, and shape: 
Are there concentrations, gaps, or outliers: 
Additional insights gained: 
Additional questions raised: 
Adjustments to be made (if any): 

## Potential Variables to be used in Model
Continuous Variables:
usd_goal
backers_count
mean_pledge

Categorical Variables:
country 
staff_pick
category