---
title: "R Notebook"
output: html_notebook
---

#############################################################
# DATA RESHAPING WITH TIDYR
#
# How to ...
#   - reshape data wide to long:    spread()
#   - reshape data long to wide:    gather()
#   - split a variable:             separate()
#   - combine two varibale values:  unite()
#
# Also: more examples at the bottom
#   - specifically, examples relevant to machine learning
#
#
# © Copyright 2017 Sharp Sight
# All rights reserved
# sharpsightlabs.com
#
#############################################################



#----------------
# LOAD PACKAGE
#----------------
```{r}
#library(tidyr)
library(startingDataScience)
```

#################
# tidyr functions
#################



#=================================
# GATHER
# - gather() converts wide to long
#
#=================================
```{r}
# inspect
head(revenue_2_untidy)
```

#-----------------------
# gather()
# - reshape wide to long
# - new key: quarter
# - new value: income
#-----------------------
```{r}
gather(revenue_2_untidy, quarter, income, -region)
```



#===================================
# SPREAD
# - spread() converts long to wide
#
#===================================
```{r}
# inspect
print(revenue_tidy)
```

#-------------------------------------------------------------------
# spread()
# - reshape long to wide
# - "quarter" variable holds the values that will become new columns
#-------------------------------------------------------------------
```{r}
spread(revenue_tidy , quarter, revenue)
```



#====================================================
# SEPARATE
# - separate() divides a column into multiple columns
#
#====================================================
```{r}
# inspect
print(founders)
```

#-------------------------------------------------
# separate()
# - break "name" into "first_name" and "last_name"
#-------------------------------------------------
```{r}
separate(founders , name , c("first_name","last_name") , sep = " ")
```



#====================================================
# UNITE
# - unite() combines multiple columns into one column
#
#====================================================

```{r}
# inspect
print(founders_2)
```

#--------------------------------------------------
# unite()
# - here we are 'uniting' first_name and last_name
#   into a single variable called 'name'
#--------------------------------------------------

```{r}
unite(founders_2, name, first_name, last_name, sep = " ")
```



################
# MORE EXAMPLES
################


#====================================================
# EXAMPLE:
#  - Boston dataset
#  - 'Boston' is a good machine learning dataset
#    and what we're doing here is typical ML
#    data exploration
#  - we're going to create a density plot of each
#    variable (all in the same plot)
#  - to do this, we'll need to reshape
#====================================================

```{r}
# get data
data(Boston, package = "MASS")
```

```{r}
# inspect
head(Boston) # print first few rows
ncol(Boston) # calculate number of columns
```


#----------------------------------------------------
# gather()
# - reshape wide to long
# - we want all of the column names in Boston to 
#   be values in a new column of our reshaped data
#----------------------------------------------------

df.boston_gathered <- gather(Boston)

# inspect
head(df.boston_gathered)


#-----------------------------------
# PLOT density plot of each variable
#-----------------------------------

ggplot(data = df.boston_gathered, aes(x = value)) +
  geom_density() +
  facet_wrap(~ key, scales = "free")



#----------------------------------------------------------------
# REDO using dplyr pipes:
# - here, we'll redo the process of reshaping and plotting Boston
#   ... we'll use dplyr pipes to do it all-in-one
#----------------------------------------------------------------

Boston %>%
  gather() %>%
  ggplot(aes(x = value)) +
    geom_density() +
    facet_wrap(~ key, scales = "free")




#======================================================================
# EXAMPLE:
# - bbbDescr dataset from 'caret'
# - note: again, this is a machine learning dataset and
#         what we're doing here is typical of ML pre-modeling work
# - ultimately, we want to create density plots of all of our variables
# - we need to reshape to get the data in a form that ggplot can use
#   in a small-multiple chart (i.e., faceting)
#======================================================================


#------------------------------
# get data from 'caret' package
#------------------------------

```{r}
library(caret)
data(BloodBrain)
```

# inspect
head(bbbDescr)  # print top few rows
dim(bbbDescr)   # get number of rows & columns ... that's a lot!



#-----------------------------------------------------
# gather()
# - reshape wide to long
# - we want to have our variable names along the rows
#-----------------------------------------------------

gather(bbbDescr) %>% head()


#--------------------------------------------------------------
# PLOT small multiple
# - we're plotting a density plot of each separate variable
# - by using dplyr pipes in this "waterfall" pattern,
#   we don't need to create and save an intermediate, gathered
#   dataframe
#--------------------------------------------------------------

bbbDescr %>%
  gather %>%
  ggplot(aes(x = value)) +
    geom_density() +
    facet_wrap(~ key, scale = "free")




# EOF